Title here

### 2017-01-24 portfolio/text/anderson-etal-rev-pfx.tex
Incorporate edits

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
S.C.A., TODO designed the study; S.C.A., TODO analyzed the
data and wrote the paper with input from all authors.}
\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sandrsn@uw.edu}

@@

% 250:
\begin{abstract}
Individuals relying on the extraction of natural resources for their
livelihood---fishers, farmers, miners, and loggers, among others---face high
levels of income variability driven by a mix of environmental, biological,
management, and
economic dynamics. A key component to managing these industries is identifying
how regulatory actions and individual behaviour affect income variability and by
extension economic stability and the sustainable use of natural resources. For
commercial fisheries, communities and vessels fishing a greater diversity of
species have less revenue variability than those fishing a lower diversity.
However, it is unclear if these benefits extend to the actions of individual
fishers and how year-to-year changes in diversification affect revenue and
revenue variability. Here, we identify three axes by which fishers in Alaska,
one of the world's most productive fishing regions, can diversify fishing
activities. We show that, despite increasing specialization over the last 30
years, fishing permits with higher species diversity reduces individual revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of existing permits has a fishery-dependent effect on revenue and is
usually (93\% probability) associated with increased revenue uncertainty
the following year. Our results demonstrate that the most effective option for
individuals to decrease revenue variability is to purchase permits to
participate in additional or more diverse fisheries. However, this option is
expensive, often limited by regulations such as catch share programs, and
consequently unavailable to many individuals. With increasing environmental
variability in a changing climate, it will be particularly important that
individuals relying on natural resources for their livelihood have effective
strategies to reduce income variability.
 % (`r g2_dens`% probability)
\end{abstract}

>>>>

% Please include corresponding author, author contribution and author declaration information
\authorcontributions{
S.C.A, E.J.W, and A.O.S designed the study with contributions from all authors;
J.C.S. also provided data;
S.C.A. and E.J.W. analyzed the data and wrote the paper with input from all
authors.}
\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sandrsn@uw.edu}

@@

% 250:
\begin{abstract}
Individuals relying on the extraction of natural resources for their
livelihood---fishers, farmers, miners, and loggers, among others---face high
levels of income variability driven by a mix of environmental, biological,
management, and
economic factors. Key to managing these industries is identifying
how regulatory actions and individual behaviour affect income variability and, by
extension, economic stability and the sustainable use of natural resources. In
commercial fisheries, communities and vessels fishing a greater diversity of
species have less revenue variability than those fishing fewer species.
However, it is unclear if these benefits extend to the actions of individual
fishers and how year-to-year changes in diversification affect revenue and
revenue variability. Here, we evaluate two axes by which fishers in Alaska,
can diversify fishing
activities. We show that, despite increasing specialization over the last 30
years, fishing a set of permits with higher species diversity reduces individual revenue
variability and fishing an additional permit is associated with higher revenue
and lower variability. However, increasing species diversity within the
constraints of existing permits has a fishery-dependent effect on revenue and is
usually (93\% probability) associated with increased revenue uncertainty
the following year. Our results demonstrate that the most effective option for
individuals to decrease revenue variability is to purchase permits to
participate in additional or more diverse fisheries. However, this option is
expensive, often limited by regulations such as license limitation or catch share programs, and
consequently unavailable to many individuals. With increasing environmental
variability in a changing climate, it will be particularly important that
individuals relying on natural resources for their livelihood have effective
strategies to reduce income variability.
 % (`r g2_dens`% probability)
\end{abstract}

### 2017-01-24 portfolio/text/ms.Rmd
Incorporate edits

\dropcap{M}anaging natural resources and sustaining the livelihoods of those that depend
on their extraction is hard. These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

@@

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in multiple areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, otherwise well-intentioned regulations may limit
how individuals diversify, or incentivize against diversification. For example,
in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, external market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Commercial fishers present an interesting example of diversification, given the
extreme risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as permitted by a license, own multiple
licenses, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies suggest
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g. counties or communities; [@kasperski2013;
@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
between diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

@@

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
countries have passed --- fishers must own a license to participate in virtually
all fisheries [@hilborn2003a]. While such regulations have helped reduce
overfishing and in many fisheries have increased profits, they limit the
available strategies to reduce income variability. For example, many permits
limit fishing to a single species, diversifying by purchasing additional fishing
permits can be expensive[^1], and some fisheries limit the entry of new
participants through individual fishing quotas (IFQs) [@grimm2012].

@@

Using a unique dataset collected over 30 years on individual fishers, rather
than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals we can track
people that switch fishery permits over time or use multiple vessels. Similarly,
we can disentangle fishers that share the same vessel. We develop and apply
a hierarchical Bayesian model to understand the effect of species diversity
within fisheries catches on individual revenue and revenue variability. Our
model builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
community-level [@sethi2014] variability. We develop a variance-function
regression model [@davidian1987; @hall1989], to simultaneously model changes in
the mean and variance of individual revenue as a function of covariates. We
simultaneously estimate the effect of species diversity (the inverse of
Simpson's diversity index [@simpson1949; @jost2006]) on individual revenue and
revenue variability, and estimate effects both within and across fishing
strategies while controlling for fishing effort and latent time effects (e.g.
changing environmental conditions). To our knowledge, this is the first
application of this type of model in any natural resource field.

@@

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold a single permit, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple-permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of this revenue declined somewhat from 1985--2000 but has since
remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
by an increase in salmon specialists in the 1980s and 1990s --- the trend
disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

Individual revenue variability decreases exponentially with higher mean fished
species diversity across combinations of state and federal fishing permits
("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
one unit higher mean species diversity had on average `r h1[[2]]`% (`r
h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern looks
qualitatively similar with or without individuals targeting salmon. As examples,
variability was highest for the herring roe gillnet and king crab fisheries.
Most strategies with the lowest estimated revenue variability combined halibut
with a salmon permit and had high levels of fished species diversity (Fig.
\ref{fig:across}).

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given set of permits is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
annual average revenue for fishers with the same set of permits, specializing
tends to be associated with increased revenue but also increased variability
(Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

@@

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a set of permits tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
S\figmirror). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing strategies is
a contributor to increased variability of revenue.

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1994, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects involving halibut and sablefish permits to vary
separately before and after the introduction of IFQs. The introduction of IFQs
reduced variability in revenue by ~25% (TODO XX--XX 95% CI) for individuals with
only halibut or sablefish permits (Fig. S\figifq A). Furthermore, for most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\figifq B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\figifq B).

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
@edwards2004]. However, in practice, the ability of individuals to effectively
diversify is limited by regulation and market forces. Here we document the
effect of individual-level changes in fished species diversity on revenue and
revenue variability across three scales. We find that (1) individuals with sets
of permits that allow high levels of species diversity tend to have lower
revenue variability. (2) Adding a fisheries permit usually increases revenue and
decreases revenue variability. However, (3) within the constraints of a permit
and accounting for effort, specializing tends to increase revenue while both
specializing and generalizing tend to correlate with greater revenue variability
with considerable differences amongst fisheries. Thus, the most effective
mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is expensive,
limited by regulations, and consequently unavailable to many individuals.

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
findings within strategies, with respect to individual year-to-year behaviour,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, local knowledge and experience are
a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
fishing patterns may sometimes be associated with a learning or adaptation phase
as individuals become accustomed to new locations, gear, or species.
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers were forced to target other species or landed
more bycatch, adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our large-scale model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behaviour [@hilborn1985; @hilborn1985a;
@branch2006] and therefore not included in our large-scale model, that might
make a specific individual's revenue less predictable if species diversity
changes.

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. We expect future research examining diversification over
finer scales will uncover additional interesting relationships. Furthermore, in
addressing the overall pattern, we have averaged out several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [@anderson2016].

With increasing environmental variability and extreme events in a changing
climate [@easterling2000], it is particularly important that individuals who
rely on natural resources as a primary source of income, such as fishers,
farmers, and forestry workers, have effective strategies to reduce risk. While
diversification can provide a powerful form of risk buffering, we have shown
that the relationship between individual changes to diversification and income
variability is complex and context specific. Furthermore, external market forces
can interact with management regulations, which have other considerable
benefits, to restrict individual diversification opportunities and promote
specialization. Although variability in individual revenue may not be actively
incorporated into management decisions, analyses such as ours help to understand
the range of diversification strategies available, trade-offs inherent in such
management decisions, and how these trade-offs may change over time.

>>>>

\dropcap{M}anaging natural resources and sustaining the livelihoods of those that depend
on their extraction is challenged by multiple sources of variability in the coupled human-natural system. These livelihoods tend to have high annual
variability in income relative to other professions [@mishra2002; @sethi2012b].
In addition to economic risk, such as changes in demand or prices, individuals
dependent on natural resources are also subject to biological and environmental
variability [@smith1986]. For example, drought and flooding are a major source
of risk for agricultural food security and farmers' incomes [@gregory2005], and
catastrophic disease outbreaks and wildfires increase risk for the logging
industry [@kirilenko2007].

@@

Individuals who rely on natural resources for income develop strategies to
reduce income variability. For example, farmers may rotate their crops to buffer
short-term climate variation and long-term environmental trends [@zentner1988].
Similarly, loggers maintain a diverse mix of tree species, age structure, and
harvest in multiple areas to buffer against disease outbreak, wildfires, and
price variation (REF). However, otherwise well-intentioned regulations may limit
how individuals diversify, or incentivize against diversification. For example,
in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
instances, market forces (e.g. prices or demand) may discourage the
adoption of diverse strategies and help promote specialization [@bradshaw2004].

Diversification may benefit commercial fishers, given the
extreme economic risk that some individuals experience, particularly relative to
terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
available to terrestrial farmers (e.g. crop insurance programs) are not
obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
ability to lower risk by diversifying their fishing portfolio. They can, for
example, target as many species as allowed by a permit, own multiple
permits, or diversify spatially and fish in multiple regions. The effectiveness
of these diversification strategies for increasing revenue and reducing revenue
variability remains uncertain. For example, theory and empirical studies demonstrate that
there can be substantial benefits to specialization for both farmers and fishers
[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
examining the effects of diversification on revenue do so at aggregate levels
that combine individuals (e.g., counties or communities; [@kasperski2013;
@sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship
among diversity, fishing revenue, and fishing revenue variability at the level
of the individual.

@@

The state and federal fisheries of Alaska provide a unique opportunity to
understand the role of individuals' strategies in determining income and income
variability. Fisheries in Alaska are among the highest-valued fisheries in the
world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
countries have passed --- fishers must own a permit to participate in virtually
all fisheries [@hilborn2003a]. While such regulations have helped reduce
overfishing and in many fisheries have increased profits, they limit the
available strategies to reduce income variability. For example, many permits
limit fishing to a single species, diversifying by purchasing additional fishing
permits can be expensive[^1], and some fisheries limit the entry of new
participants through individual fishing quotas (IFQs) [@grimm2012].

@@

<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth a footnote to address the potential critique that at the indivdual level we don’t observe their wages in non-fishing activities.  I’m not sure if this is worth bringing up, but this is always the case anyway with vessels where the skipper may do other things. -->

Using a unique dataset collected over 30 years on individual fishers, rather
than fleets or vessels, we examine how revenue and revenue variability of
individuals is affected by diversification. By tracking individuals, it is possible to track
people that switch fishery permits over time or use multiple vessels and differentiate fishers that share the same vessel. We developped and applied
a hierarchical Bayesian model to understand the effect of species diversity
within fisheries catches on individual revenue and revenue variability. Our
model builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
community-level [@sethi2014; @cline2017] variability. We developped a variance-function
regression model [@davidian1987; @hall1989], to simultaneously model changes in
the mean and variance of individual revenue as a function of covariates. We
simultaneously estimate the effect of species diversity (the inverse of
Simpson's diversity index [@simpson1949; @jost2006]) on individual revenue and
revenue variability, and estimate effects both within and across fishing
strategies while controlling for fishing effort and latent time effects (e.g.
changing environmental conditions). The model was used to evaluate two axes by which fishers in Alaska can diversify fishing activities: (1) choice of permit strategy (i.e., permit or set of permits fished), and (2) choice of what species to fish *within* a permit strategy. To our knowledge, this is the first
application of this type of model in any natural resource field.

@@

<!-- TODO AH: do we need to make a comment on the early 80s data since  we mention 1980 here but say that we starting in 1985? If I were reading,  I would wonder if you have the data, why don’t you use it from the start? on page 15 you say 1986 -->

We evaluated trends in revenue and diversification over time for individuals
fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
majority (70--90%) hold a single permit, with a slight increase in single-permit
holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
halibut or salmon permit holders --- compared to other fisheries, halibut permit
holders are more likely to hold halibut plus other permits, whereas salmon
permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
Despite making up only 10--30% of the individuals, multiple-permit holders earn
approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
diversity of the catch that generated this revenue declined somewhat from 1985--2000 but has since
remained steady (Fig. \ref{fig:ts}D). This trend appears largely driven
by an increase in salmon specialists in the 1980s and 1990s, disappearing
if salmon license holders are excluded (Fig. \ref{fig:ts}D).

@@

<!-- TODO AB: To me it is more intuitive that the darker, hotter colors would represent higher values. Consdier reversing the color scale to go from red (highest # of permit holders) to yellow (lowest #).  -->

Individual revenue variability decreases exponentially with higher mean fished
species diversity across combinations of state and federal fishing permits
("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
one unit higher mean species diversity had on average `r h1[[2]]`% (`r
h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern is
qualitatively similar with or without individuals targeting salmon. As examples,
revenue variability was highest for the herring roe gillnet and king crab fisheries.
Most strategies with the lowest estimated revenue variability combined halibut
with a salmon permit and had high levels of fished species diversity (Fig.
\ref{fig:across}).

@@

In contrast to the clear across-strategy effects of species diversity on revenue
and revenue variability, the effect of changing fished species diversity for an
individual *within* the constraints of a given strategy (set of permits) is more variable
(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
annual average revenue for fishers with the same set of permits, specializing
tends to be associated with increased revenue but also increased variability
(Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
decrease in species diversity, and across strategies weighted equally, `r
g2_dens`% of the probability density supports a positive relationship between
specialization and revenue.

@@

Changing (increasing or decreasing) an individual's fished species diversity
within the constraints of a set of permits tends to be associated with greater
expected variability or uncertainty in year-to-year revenue (Fig.
\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
S\figmirror). This effect was strongest for individuals with a halibut
permit or halibut and sablefish permits. Individuals with these strategies
experienced a ~2% increase in variability for a 1% change in the year-to-year
ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
species diversity also leads to increased revenue variability (Fig.
S\ref{fig:mirror}A), suggesting that the act of changing the number of species fished for a given set of permits is
a contributor to increased variability of revenue.

@@

Alaskan halibut and sablefish fisheries were converted to individual fishery
quota (IFQ) management in 1995, which potentially confounds our estimates with
a major regulatory change. Therefore, we refit our models allowing all
strategy-level effects involving halibut and sablefish permits to vary
separately before and after the introduction of IFQs. The introduction of IFQs
was associated with a reduction in revenue variability of ~25% (TODO XX--XX 95% CI) for individuals with
only halibut or sablefish permits (Fig. S\figifq A).
<!-- TODO AH: We need to note that the populatoin changed -->
Furthermore, for individuals within most
strategies, the association between changing species diversity and increased
revenue variability became stronger after the introduction of IFQs (Fig.
S\figifq B). This was especially true for individuals fishing only halibut
or sablefish and less so for individuals combining halibut or sablefish with
a salmon permit (Fig. S\figifq B).

@@

For individuals deriving livelihoods from natural resources, theory suggests
that adopting strategies that exploit a diversity of assets provides substantial
benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
@edwards2004]. However, in practice, the ability of individuals to effectively
diversify is limited by regulation and market forces. Here we document the
effect of individual-level changes in fished species diversity on revenue and
revenue variability across two scales. We find that individuals with sets
of permits that allow high levels of species diversity tend to have lower
revenue variability and adding a fisheries permit usually increases revenue and
decreases revenue variability. However, within the constraints of a permit strategy
and accounting for effort, specializing tends to increase revenue while changes to fished species diversity tend to correlate with greater revenue variability
although there are considerable differences amongst fisheries. Thus, the most effective
mechanism by which species diversity can lower revenue variability for
individual fishers --- switching or adding fishing permits --- is expensive,
limited by regulations, and consequently unavailable to many individuals.

Our findings across fishing strategies are in line with previous research at the
community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
findings within strategies, with respect to individual year-to-year behavior,
provide new insights. Here, specializing tended to be associated with increased
revenue. This could be for a number of reasons. First, specialization can
increase fishing efficiency [@smith1986]. Efficiency gains from specialization
can be seen in other natural resource sectors. For example, rice farmers in
Ghana can increase their efficiency by devoting more land and labor to rice
production compared to other crops [@abdulai2000]. Second, given the
availability of valuable resources, specializing can be highly profitable. For
instance, when cocoa prices have been high, Nigerian farmers have tended to
specialize on the crop [@mustapha1999] and specializing on valuable tobacco
farming can be the optimal strategy regardless of alternative crops
[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
valuable sockeye salmon benefited the most from specializing. However,
specializing can have longterm risks if market, management, biological, or
environmental conditions change. For example, U.S. farmers specialized on wheat,
corn, or soybeans were hit harder than diversified farmers in the early 1980s
when a financial crisis hit [@ikerd1990].
<!-- TODO AB: We might want to include more examples / references from fisheries. -->
Third, diversifying --- a lack of
specialization --- can happen as a consequence of non-ideal circumstances such
as poor availability of a more valuable resource. For example, Ugandan
households relying on forestry resources for income increase their income
portfolio diversity following downward economic shocks [@debela2012].
Because specialization inherently holds risk from ``putting all one's eggs in one
basket,'' it is not suprising that a decline in the most valuable resource
would lead to a shift towards diversification.

We also found that, within strategies, altering species diversity tended to be
associated with increased revenue variability. There are number of possible
reasons for this relationship. For instance, local knowledge and experience are
a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
fishing patterns may sometimes be associated with a learning or adaptation phase
as individuals become accustomed to new locations, gear, or species.
<!-- TODO AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies. Otherwise, statements like this one aren’t really anchored to anything concrete.  For example, what does increased specialization look like in the Hal ll permit, which showed one of the greatest effects of specialization on revenue diversity? How many permits could a Hal ll permit holder potentially fish? This goes back to my comment re: Table 1 (see p. 15) -->
Additionally, changes in diversity might reflect years in which targeted species
were less available and fishers instead chose to target other species or landed
more bycatch,
<!-- TODO AB: This is the first time bycatch is explicitly mentioned in the study and it raises some questions about how target vs. bycatch species are defined and considered in the analysis. Some clarification in the methods would be helpful. -->
adding stochasticity to their revenue. Most importantly, this
increased variability reflects an inability for our model to predict
an individual's revenue in a subsequent year. There are many reasons, often
unique to an individual permit holder's behavior [@hilborn1985; @hilborn1985a;
@branch2006] and therefore not in our model, that might
make a specific individual's revenue less predictable if species diversity
changes.

@@

Our analysis provides a synthetic view of the effects of species diversity on
individuals over 30 years in one of the most productive fishing regions in the
world. However, this large temporal and geographical scope potentially obscures
time or regional-variation in the consequences of diversification on revenue and
revenue variability. Furthermore, in
addressing the overall pattern, we have averaged out the potential signal of several major discrete
events, such as the Exxon Valdez oil spill in 1989, that had significant
consequences for the ecological and social systems in Alaska [@paine1996]. In
addition to revenue or income variability, it is important to understand the
benefits of diversification in the context of rare and extreme disturbances
which occur in natural systems worldwide [@anderson2017].
<!-- TODO AH: Isn’t the variabiity you mention in this sentence part of this? -->

With increasing environmental variability and extreme events in a changing
climate [@easterling2000], it is particularly important that individuals who
rely on natural resources as a primary source of income, such as fishers,
farmers, and forestry workers, have effective strategies to reduce risk. While
diversification can provide a powerful form of risk buffering, we have shown
that the relationship between individual changes to diversification and income
variability is complex and context specific. Furthermore, external market forces
can interact with management regulations, which have other considerable
benefits, to restrict individual diversification opportunities and promote
specialization. Although variability in individual revenue may not be actively
incorporated into management decisions, analyses such as ours help to explain
the range of diversification strategies available, trade-offs inherent in such
management decisions, and how these trade-offs may change over time.
